Performance for the masses: Experiments with A Web Based Architecture to Harness Volunteer Resources for Low Cost Distributed Evolutionary Computation

### 2016-01-26 GECCO/volunteer-clouds-gecco16.tex

%Mario: Is NodIO presented here after the cut? 
%TO DO: Add reference
There is, then, space for frameworks that do not make any of these
assumptions, such as {\sf NodIO} presented in \cite{2016arXiv160101607Manom}. {\sf NodIO} is basically 
%assumptions, such as the one presented in this paper, which will be
% named {\sf NodIO}. {\sf NodIO} is basically
% (Paloma) Is the "first" word really needed? 
% (Mario) I guess is more related to fundamentally or primarily but "basically" sounds better
a two tier client/server 
application, in which clients can run on a browser. 
Therefore, they can be embedded in any web-enabled device, including many over which 
developers have no control.

@@

The main problem with Java and all the other browser-embedded
technologies is that they are not {\em universal} in the
sense that an extra component, namely, the Java virtual machine or another
virtual environment has to be installed in the browser.
Also these virtual environments have become targets of malicious attacks; 
for instance numerous vulnerabilities have been discovered in 
the proprietary Adobe Flash runtime environment and
video player \cite{ford2009analyzing,watanabe2010new}. The need for
these additional environments has been mitigated with the
adoption of the HTML5 standard \cite{anthes2012html5}.
JavaScript \cite{flanagan2006javascript}, however, has become a set of standards
\cite{ECMA-262} that include the language and its components. Since
early on, it was adopted by the industry and also by scientists,
who used them for creating an evolutionary algorithm (EA from now on) on
the browser as early as 1998 \cite{jj-ppsn98}.

JavaScript can be used either for unwitting
\cite{unwitting-ec} or volunteer 
\cite{langdon:2005:metas,gecco07:workshop:dcor} distributed
evolutionary computation and it has been used ever since by several
authors, including more recent efforts
\cite{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210} 
that even
used the client's GPU \cite{duda2013gpu} or create an object-oriented
framework for evolution in the browser \cite{EvoStar2014:jsEO}. Many other researchers have
used Java \cite{chong:1999:jDGPi} and others have gone away from the
server-based paradigm to embrace peer to peer systems
\cite{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These computing
platforms avoid single points of failure (the server) but, since they
need a certain amount of infrastructure installed to start, the
threshold to join them is much lower. % (Paloma) Do yo mean... higher? A low threshold means that is easy to happen something.

Recent works propose the use of cloud computing services for the components of
client-server distributed EAs. Cloud resources
can reduce operational costs by outsourcing both hardware and software maintenance
to the cloud provider. Another advantage is that they enable the provisioning of computing resources beyond what
is available in most laboratories, allowing researchers to
scale algorithms at reasonable costs. Researchers have proposed the use of
public clouds such as Amazon EC2 \cite{CloudScale}, Google App Engine\cite{di2013towards},
and DropBox~\cite{mericloud}. The use of JavaScript to implement
server side components
gives software developers certain advantages in the cloud
paradigm, since it is a first-class citizen
in most Platforms as a Service (PaaS) products as the {\tt Node.js} framework is supported by most of them \cite{wood13:nodejs:paas}. Many web services
have native JavaScript libraries too, and also use the JSON (JavaScript Object Notation) format for data interchange. Another advantage is that
developers can use the same software tools for both client and server
side development.

JavaScript is also compatible with the functional programming paradigm \cite{Cousineau1998,MacLennan1990,Thompson1996}.
According to \cite{swanresearch2015}, the benefit of using this kind of paradigm in
metaheuristics optimization is that it allows an improvement in communicability,
reproductibility, interoperability, automated assembly, knowledge engineering,
and efficiency.

In the work presented in this paper, JavaScript has been used throughout the full
stack, which has the advantage of using the self-same lines of code for
writing an evolutionary algorithm for the
desktop or command line and browser clients, both based in the EC
library {\sf NodEO} \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time 
when writing code and leverages 
the expertise acquired
% (Paloma) Not sure if "acquired" should be written before "expertise"... sounds better to me, anyways
 in the library itself.

@@

On the other hand, measuring the performance of the resulting metacomputer
involves understanding the dynamics of this kind of systems. Initial
work was done for peer to peer systems by Stutzbach et
al. \cite{stutzbach2006understanding} and extended to volunteer
computing by Laredo et al. \cite{churn08,laredo2008rcp}. 
%But the raw material of           
%volunteer computing, % (Paloma) Whith regard to?  Maybe it was colloquial/analogy
But some of the essential metrics in volunteer computing like the
number of users or the time spent by every one in the
computation in browser-based volunteer computing experiments, have
only been studied in a limited way in 
\cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
run. Studies using volunteer computing platforms such as SETI@home
\cite{javadi2009mining} found out that the Weibull, log-normal, and
Gamma distribution 
modeled quite well the availability of resources in several clusters
of that framework; the shape of those distributions is a skewed bell
with more resources in the {\em low} areas than in the high areas:
there are many users that give a small amount of cycles, while there
are just a few that give many cycles. This is in concordance with the
results obtained in \cite{agajaj}. Although SETI@home works
at a larger scale in both quantity of volunteers and the scope of projects, 
our initial tests indicate that
this kind of distribution might be the most appropriate to describe
user participation in several different metrics. However, in this
paper we will perform different experiments in different setups and
will try to come up with a more precise model. Next we describe the
framework itself.

As far as we know, this paper presents one of the few experiments using % (Paloma) Change to "which" in order to not repeating "that" all the time
computational resources that are as dissimilar as smartphones and
powerful laptops 
or workstation computers in a research center. 
%The methodology used
% (Paloma) I would say "the used methodology"
The algorithms used, as well as the methodology 
for gathering resources and baseline results that we will use for
comparison have been described in
\cite{2016arXiv160101607Manom}. We will focus next on the experimental
results using the first version of the framework.

@@

>>>>

%Mario: Is NodIO presented here after the cut? 
% Not sure. Maybe not. - JJ
%TO DO: Add reference
There is, then, space for frameworks that do not make any of these
assumptions, such as {\sf NodIO} presented in \cite{2016arXiv160101607Manom}. {\sf NodIO} is basically 
%assumptions, such as the one presented in this paper, which will be
% named {\sf NodIO}. {\sf NodIO} is basically
% (Paloma) Is the "first" word really needed? 
% (Mario) I guess is more related to fundamentally or primarily but "basically" sounds better
a two tier client/server 
application, in which clients can run on a browser. 
Therefore, they can be embedded in any web-enabled device, including many over which 
developers have no control.

@@

The main problem with using Java and all the other browser-embedded
technologies is that they are not {\em universal} in the
sense that an extra component, namely, the Java virtual machine or another
virtual environment has to be installed in the browser.
Also these virtual environments have become targets of malicious attacks; 
for instance numerous vulnerabilities have been discovered in 
the proprietary Adobe Flash runtime environment and
video player \cite{ford2009analyzing,watanabe2010new}. The need for
these additional environments has been mitigated with the
adoption of the HTML5 standard \cite{anthes2012html5}.
JavaScript however, has become a set of standards
\cite{ECMA-262} that include the language and its components.

JavaScript can be used either for unwitting, that is, stealth,
\cite{unwitting-ec} or volunteer 
\cite{langdon:2005:metas,gecco07:workshop:dcor} distributed
evolutionary computation and it has been used ever since by several
authors, including more recent efforts
\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210} 
that even
used the client's GPU \cite{duda2013gpu}. Even if, as said above, Java
is not the best option, many other researchers have
used Java \cite{chong:1999:jDGPi} and others have gone away from the
server-based paradigm to embrace peer to peer systems
\cite{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These
computing 
platforms avoid single points of failure (the server) but, since they
need a certain amount of infrastructure installed to start, the
threshold to join them is much higher. % (Paloma) Do yo
                                % mean... higher? A low threshold
                                % means that is easy to happen
                                % something.
%Correct. But maybe in this new version we should eliminate this
%paragraph altogether. 
% The next paragraph has been suppresed.

JavaScript is also compatible with the functional programming paradigm \cite{Cousineau1998,MacLennan1990,Thompson1996}.
According to \cite{swanresearch2015}, the benefit of using this kind of paradigm in
metaheuristics optimization is that it allows an improvement in communicability,
reproductibility, interoperability, automated assembly, knowledge engineering,
and efficiency. In the work presented in this paper, JavaScript has been used throughout the full
stack, which has the advantage of using the self-same lines of code for
writing an evolutionary algorithm for the
desktop or command line and browser clients, both based in the EC
library {\sf NodEO} \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time 
when writing code and leverages 
the expertise acquired
% (Paloma) Not sure if "acquired" should be written before
% "expertise"... sounds better to me, anyways
% Don't know what you mean. Write it down and we'll see - JJ
 in the library itself.

@@

On the other hand, measuring the performance of the resulting metacomputer
involves understanding its dynamics. Initial
work was done for peer to peer systems by Stutzbach et
al. \cite{stutzbach2006understanding} and extended to volunteer
computing by Laredo et al. \cite{churn08,laredo2008rcp}. 
%But the raw material of           
%volunteer computing, % (Paloma) Whith regard to?  Maybe it was colloquial/analogy
But some of the essential metrics in volunteer computing like the
number of users or the time spent by every one in the
computation in browser-based volunteer computing experiments, have
only been studied in a limited way in 
\cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
run. Studies using volunteer computing platforms such as SETI@home
\cite{javadi2009mining} found out that the Weibull, log-normal, and
Gamma distribution 
modeled quite well the availability of resources in several clusters
of that framework; the shape of those distributions is a skewed bell
with more resources in the {\em low} areas than in the high areas:
there are many users that give a small amount of cycles, while there
are just a few that give many cycles. This is in concordance with the
results obtained in \cite{agajaj}. Although SETI@home works
at a larger scale in both quantity of volunteers and the scope of projects, 
our initial tests indicate that
this kind of distribution might be the most appropriate to describe
user participation in several different metrics. However, in this
paper we will perform different experiments in different setups and
will try to come up with a more precise model. Next we describe the
framework itself.

As far as we know, this paper presents one of the few experimental
results using a volunteer computing system that uses  % (Paloma) Change to "which" in order to not repeating "that" all the time
computational resources that are as dissimilar as smartphones and
powerful laptops 
or workstation computers in a research center. This will be described next.

@@

% A bit of description of the framework, with reference to published
% reports - JJ
